Evaluation of classical correlation functions from 2/3D images on CPU and GPU architectures: introducing CorrelationFunctions.jl

### 2023-09-06 paper-correlations.tex
Наш ответ на критику

\section{Computational methodology}
\code{CorrelationFunctions.jl} is capable to compute all aforementioned
correlation functions. Where possible we provide two implementations for each
CF. The first implementation computes a correlation function in specific
predefined directions (i.e., orientation of test vectors thrown into a sample is
fixed) \cite{jiao2014chawla,EPL1}.  There are one predefined direction in 1D
case, four directions in 2D case (two orthogonal and two diagonal) and thirteen
directions in 3D case (three orthogonal, six diagonal within each orthogonal
plane and four diagonal). All directions available are schematically represented
in \cref{fig:directions}. The line segment scanning approach is slower for some
CFs but requires much less memory resources. The full map approach establishes
correlations between all points on the image is usually faster, as it leverages
on Fast Fourier Transform computational speed.  Both implementations are
contained in \code{Directional} and \code{Map} modules of the package,
correspondingly.

@@

For surface-surface and surface-void functions we use an edge detection filter
\cite{samarin2023robust} and then calculate either autocorrelation of the edge
or cross-correlation of the edge and the void phase. The edge is extracted by
convolving an input with a short high-pass filter $H$. It has a width $7$ and
all its coefficients with exception of the central coefficient are inversely
proportional to distance from the center:
\begin{equation}
  \begin{aligned}
    H_{ij} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2}} & \quad i \ne 3, j \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right. \\
    H_{ijk} &= S \left\{
    \begin{array}{cc}
      \frac{1}{\sqrt{(i-3)^2 + (j-3)^2 + (k-3)^3}} & \quad i \ne 3, j \ne 3, k
      \ne 3 \\
      C & \quad \text{otherwise}
    \end{array}
    \right.
  \end{aligned}
  \label{eq:filter-7x7}
\end{equation}
where $C$ is such that all coefficients of $H$ sum to zero, and $S$ equals to
$30.45849$ in 2D and $172.96232$ in 3D case. The parameters of the filter were adjusted
based on comparison against exact computations of surface functions on smooth-boundary
sets \cite{postcov23}.

@@

\section{Summary}
\label{sec:summary}
In this paper we introduced \code{CorrelationFunctions.jl} open-source package
developed in Julia and capable of computing all classical correlation functions
based on 2D and 3D input images. Our code is capable of evaluating two-point
probability, cluster, lineal-path, surface-surface, surface-void, pore-size and
chord-length functions on both CPU and GPU architectures. There possible, we
presented two types of computations: full correlation map (correlations of each
point with other pints on the image, that also allows obtaining ensemble
averaged CF) and directional correlation functions (currently in major
orthogonal and diagonal directions). Such an implementation allowed for the
first time to assemble a completely free solution to evaluate correlation
functions under any operating system with well documented API. Our package
includes automatic tests against analytical solutions that are described in this
paper. We measured wall times for all CPU and GPU implementations and as a rule
of thumb full correlation maps on GPU are faster than directional scans using
CPU due to the usage of fast Fourier transform. However, full maps require more
RAM and, thus, are limited to available GPU RAM resources. On the other hand,
directional CFs are memory efficient and can be evaluated for huge datasets -
this way they are the first candidates for structural data compression of
feature extraction. The package itself is available through Julia package
ecosystem and on GitHub (https://github.com/fatimp/CorrelationFunctions.jl), the
latter source also contains documentation and additional helpful resources such
as tutorials.  We believe that a single powerful computational tool such as
\code{CorrelationFunctions.jl} will significantly facilitate the usage of
correlation functions in numerous areas of structural description and research,
as well as in machine learning applications.

>>>>

\section{Computational methodology}
\code{CorrelationFunctions.jl} is capable to compute all aforementioned
correlation functions. Where possible we provide two implementations for each
CF. The first implementation computes a correlation function in specific
predefined directions (i.e., orientation of test vectors thrown into a sample is
fixed) \cite{jiao2014chawla,EPL1}.  There are one predefined direction in 1D
case, four directions in 2D case (two orthogonal and two diagonal) and thirteen
directions in 3D case (three orthogonal, six diagonal within each orthogonal
plane and four diagonal). All directions available are schematically represented
in \cref{fig:directions}. The line segment scanning approach is slower for some
CFs but requires much less memory resources. The full map approach establishes
correlations between all points on the image and is faster. Both implementations
are contained in \code{Directional} and \code{Map} modules of the package,
correspondingly.

@@

For surface-surface and surface-void functions we use an edge detection filter
\cite{samarin2023robust} and then calculate either autocorrelation of the edge
or cross-correlation of the edge and the void phase. The edge is extracted by
convolving an input with a short high-pass filter $H$. It has a width $7$ and
all its coefficients with exception of the central coefficient are inversely
proportional to distance from the center:
\begin{equation}
  \begin{aligned}
    H_k &= S \left\{
    \begin{array}{ll}
      -\sum\limits_{\substack{l \\ l \ne 0}} E_l & \quad k = 0 \\
      1 / \rho(k, 0) & \quad \text{otherwise}
    \end{array}
    \right. \\
    k &\in \{-3, -2, \dots, 3\}^D
  \end{aligned}
  \label{eq:filter-7x7}
\end{equation}
where $S$ equals to $30.45849$ in 2D and $172.96232$ in 3D case. The parameters
of the filter were adjusted based on comparison against exact computations of
surface functions on smooth-boundary sets \cite{postcov23}.

@@

\section{Summary}
\label{sec:summary}
In this paper we introduced \code{CorrelationFunctions.jl} open-source package
developed in Julia and capable of computing all classical correlation functions
based on 2D and 3D input images. Our code is capable of evaluating two-point
probability, cluster, lineal-path, surface-surface, surface-void, pore-size and
chord-length functions on both CPU and GPU architectures. There possible, we
presented two types of computations: full correlation map (correlations of each
point with other pints on the image, that also allows obtaining ensemble
averaged CF) and directional correlation functions (currently in major
orthogonal and diagonal directions). Such an implementation allowed for the
first time to assemble a completely free solution to evaluate correlation
functions under any operating system with well documented API. Our package
includes automatic tests against analytical solutions that are described in this
paper. We measured wall times for all CPU and GPU implementations and as a rule
of thumb full correlation maps are faster than directional scans because
calculation of correlation maps applies fast Fourier transform to the whole
input array while, when calculating scans, FFT needs to be applied to each scan
separately. Calculation on GPU shows performance increase when compared to CPU,
as expected.  However, full maps require more RAM and, thus, are limited to
available RAM resources. On the other hand, directional CFs are memory efficient
and can be evaluated for huge datasets - this way they are the first candidates
for structural data compression of feature extraction. The package itself is
available through Julia package ecosystem and on GitHub
(https://github.com/fatimp/CorrelationFunctions.jl), the latter source also
contains documentation and additional helpful resources such as tutorials.  We
believe that a single powerful computational tool such as
\code{CorrelationFunctions.jl} will significantly facilitate the usage of
correlation functions in numerous areas of structural description and research,
as well as in machine learning applications.
